Language learning in context via recursive pragmatic reasoning

### 2013-05-31 praglearn.tex

Something
fundamentally different happens in this case than in the specificity
implicature case (\S\ref{sec:learning-specificity-implic}), however. There, we
saw that our model could hear pragmatically strengthened usages
of words while recovering the underlying literal meaning. In the Horn
case, though, the pragmatic strengthening depends essentially on
ignorance about literal meaning. As the agents gather more data,
the implicature disappears, to be replaced by a belief that
``cheap'' \textit{literally} means \textsc{common}. To demonstrate
this phenomenon, we queried each agent in each dyad about how they would refer to
or interpret each object and word, \textit{if} the two objects were
equally common, which cancels the Horn implicature. As shown in
Fig.~\ref{fig:emergence-average} (right), after 30 turns, in nearly 70\%
of dyads both $S$ and $L$ believed in the ``good'' mapping even in
this implicature-free case, while less than 20\% believed in the
``bad'' mapping (with the rest being inconsistent). This makes the
interesting prediction that while specificity implicatures can survive
indefinitely in a language system, any Horn implicatures that arise
consistently should transition to becoming part of the literal meaning
of the words involved.

>>>>

Something
fundamentally different happens in this case than in the specificity
implicature case (\S\ref{sec:learning-specificity-implic}), however. There, we
saw that our model could hear pragmatically strengthened usages
of words while recovering the underlying literal meaning. In the Horn
case, though, the pragmatic strengthening depends essentially on
ignorance about literal meaning. As the agents gather more data,
the implicature disappears, to be replaced by a belief that
``cheap'' \textit{literally} means \textsc{common} (and did all along). To demonstrate
this phenomenon, we queried each agent in each dyad about how they would refer to
or interpret each object and word, \textit{if} the two objects were
equally common, which cancels the Horn implicature. As shown in
Fig.~\ref{fig:emergence-average} (right), after 30 turns, in nearly 70\%
of dyads both $S$ and $L$ believed in the ``good'' mapping even in
this implicature-free case, while less than 20\% believed in the
``bad'' mapping (with the rest being inconsistent). Our model therefore makes the
interesting prediction that while specificity implicatures can survive
indefinitely in a language system, any Horn implicatures that arise
consistently should transition to becoming part of the literal meaning
of the words involved.
